---
title: "PredicTR analysis"
output: html_document
date: "2025-04-30"
---

## Install/load packages and import data

```{r packages, include=FALSE}

library(survival)
library(survcomp)
library(MASS)
library(rms)
library(polspline)
library(knitr)
library(readxl)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(survival,
               Hmisc,
               pec,
               timeROC,
               riskRegression,
               rms,
               knitr,
               kableExtra,
               tidyverse)

options(show.signif.stars = FALSE)  # display statistical intelligence
```

#Load data
```{r setup, include=FALSE}
#Load data

data <- read_excel("~/Downloads/Lauren/Med school/Research/PredicTR/Predictor cleaned dataset ( recieved Syed Haider 2.3.17 merged_data) +Manchester gender info.xls")

#str(data)
training <-  data[data$set == "Training", ]
val <- data[data$set == "Validation", ]

all_vars <- c("BCL2", "COX2", "CyclinD1", "EGFRext", "HIF1alpha", "PLK1","Survivin", 
              "p16", "HPVISHHR","TILS", "Gender","Age at Diagnosis", "Surgery", 
              "RT", "CT", "T", "N", "Smoking status")
cont_vars <- c("BCL2", "COX2", "CyclinD1", "EGFRext", "HIF1alpha", "PLK1","Survivin")

biomarkers <- c("BCL2", "COX2", "CyclinD1", "EGFRext", "HPVISHHR", "p16", "PLK1", "Survivin","TILS")

#Set which vars to select and keep them
predictr_data <- subset(data, select = c(all_vars, "t.os", "Outcome", "set"))

predictr_data_dx_date <- subset(data, select = c(all_vars, "t.os", "Outcome", "set", "Diagnostic Biopsy Date"))
#str(predictr_data_dx_date)

## Numeric conversion except gender and set.
to_numeric <- function(data, vars) {
  for (col in vars) {
    if (col!="Gender" & col!="set"){
      if (is.factor(data[[col]]) || is.character(data[[col]])) {
            data[[col]] <- as.character(data[[col]])
            data[[col]][data[[col]] == "NA"] <- NA
            data[[col]] <- as.numeric(data[[col]])
          }
    }

    
  }

#  #scale the continuous vars, "cont_vars" then z-transform
#  for (col in cont_vars) {
#    data[[col]] <- as.numeric(as.character(data[[col]]))
#    data[[col]] <- scale(data[[col]]/30)[,1]
# }
 
  # Dichotomise p16
#  data$p16 <- as.numeric(as.character(data$p16))
#  data$p16 <- factor(ifelse(data$p16 >= 140, 1, 0))


  return(data)
} 


#Convert everything except Gender to numeric
predictr_data <- to_numeric(predictr_data, c(all_vars, "Outcome", "t.os"))
predictr_data_dx_date <- to_numeric(predictr_data_dx_date, c(all_vars, "Outcome", "t.os", "Diagnostic Biopsy Date"))
str(predictr_data)
str(predictr_data_dx_date)
# Convert "NA" to NA in Gender and set:
predictr_data$Gender[predictr_data$Gender == "NA"] <- NA
predictr_data$set[predictr_data$set == "NA"] <- NA

predictr_data_dx_date$Gender[predictr_data_dx_date$Gender == "NA"] <- NA
predictr_data_dx_date$set[predictr_data_dx_date$set == "NA"] <- NA


#Remove NAs (based on which variables?)
rm.na0 <- predictr_data[complete.cases(predictr_data), ]
#Gives 471 observations
rm.na1 <- predictr_data[!is.na(predictr_data$t.os), ]
# Gives 835 observations

rm.na2 <- predictr_data[!is.na(predictr_data$t.os) & !is.na(predictr_data$Outcome), ]
# Gives 809 observations

# colSums(is.na(rm.na2_dx_date))
# Here, 'Age at Diagnosis' has 2 NAs: maybe that's another required var?

rm.na1_dx_date <- predictr_data_dx_date[!is.na(predictr_data_dx_date$t.os) & !is.na(predictr_data_dx_date$`Diagnostic Biopsy Date`) & !is.na(predictr_data_dx_date$`Age at Diagnosis`), ]
# Gives 833 obs, once requiring 'Age at Diagnosis'

#Convert continuous variables to z scores, factor p16, HR-HPV DNA, TILs, age, gender, T-category, N-category, smoking status, surgery, RT, CT (using lowest category as baseline group). Need to make sure everything is already numeric
preprocess <- function(data, cont_vars) {
#scale the continuous vars, "cont_vars" then z-transform
  for (col in c(cont_vars)) {
    data[[col]] <- as.numeric(as.character(data[[col]]))
    data[[col]] <- scale(data[[col]]/30)[,1]
 }
 
  # Dichotomise p16
  data$p16_binary <- factor(ifelse(data$p16 >= 140, 1, 0))
  # Dichotomise age
  data$`Age at Diagnosis` <- factor(ifelse(data$`Age at Diagnosis`>= 50, 1, 0))
  # Factor variables:
  data$HPVISHHR <- factor(data$HPVISHHR)
  data$TILS <- factor(data$TILS)
  data$T <- factor(data$T)
  data$N <- factor(data$N)
  data$`Smoking status` <- factor(data$`Smoking status`)
  data$Surgery <- factor(data$Surgery)
  data$RT <- factor(data$RT)
  data$CT <- factor(data$CT)
  data$Gender <- factor(data$Gender)
  
  return(data)
} 

data_scaled <- preprocess(rm.na1, cont_vars)

data_scaled_dx_date <- preprocess(rm.na1_dx_date, cont_vars)
str(data_scaled)
str(data_scaled_dx_date)

#Split into training and validation
training_data <- data_scaled[data_scaled$set == "Training", ]
#n=530
training_data_dx_date <- data_scaled_dx_date[data_scaled_dx_date$set == "Training",]
#n=528

val_data <-data_scaled[data_scaled$set == "Validation", ]
#n=305
val_data_dx_date <- data_scaled_dx_date[data_scaled_dx_date$set == "Validation",]
#n=305

str(training_data)
str(training_data_dx_date)



```

```{r}




```

## Table S4

```{r}
summary(predictr_data[biomarkers])
```
## Table S5
n are different - +/- 1 or 2 observations in almost all columns
```{r}
#n column
summary(training_data)[]
nrow(training_data_dx_date)
summary(training_data_dx_date)[]
col_n <- nrow(training_data_dx_date) - colSums(is.na(training_data_dx_date))
col_n
#Make sure names agree
names(training_data)
names(col_n)[names(col_n) == "Age at Diagnosis"] <- "Age.at.Diagnosis"
names(col_n)[names(col_n) == "Smoking status"] <- "Smoking.status"
names(col_n)
```

##Univariable Cox
```{r}
# Function for univariable Cox
univariable_cox <- function(data, vars, n_obs) {
  results <- lapply(vars, function(var) {

    
    formula <- as.formula(paste("Surv(t.os, Outcome) ~", var))
    model <- coxph(formula, data = data)
    coef_summary <- summary(model)$coefficients
    return(data.frame(Variable = var, 
                      Coefficient = coef_summary[, "coef"],
                      Lower = coef_summary[, "coef"] - 1.96 * coef_summary[, "se(coef)"],
                      Upper = coef_summary[, "coef"] + 1.96 * coef_summary[, "se(coef)"],
                      P = coef_summary[, "Pr(>|z|)"],
                      n = n_obs[[var]]
                      ))
  })
  results_df <- bind_rows(results)
  # Adjust p-values using Benjamini & Hochberg method
  results_df$Q <- p.adjust(results_df$P, method = "BH")
  return(bind_rows(results_df))
}

colnames(training_data)[colnames(training_data) == 'Age at Diagnosis'] <- 'Age.at.Diagnosis'
colnames(training_data)[colnames(training_data) == 'Smoking status'] <- 'Smoking.status'

all_vars1 <- c("BCL2", "COX2", "CyclinD1", "EGFRext", "HIF1alpha", "PLK1","Survivin", 
              "p16_binary", "HPVISHHR","TILS", "Gender","Age.at.Diagnosis", "Surgery", 
              "RT", "CT", "T", "N", "Smoking.status")

# Summary of key columns
str(training_data[, c("t.os", "Outcome", all_vars1)])
sapply(training_data[, c("t.os", "Outcome", all_vars1)], function(x) sum(is.na(x)))
#levels(training_data$Gender)
#F=1, M=2

print(univariable_cox(training_data, all_vars1, col_n))



```
## (Table S6): Biomarker-only original multivariable prognostic model
```{r}
# Function 5: Multivariable Cox with Backward Elimination (S6)
print_cox_model <- function(data, vars) {
  formula <- as.formula(paste("Surv(t.os, Outcome) ~", paste(vars, collapse = " + ")))
  full_model <- coxph(formula, data = data)
  k <- qchisq(0.25, 1, lower.tail = FALSE)  # ~1.323
  step_model <- step(full_model, direction = "backward", k = 2, trace = 1)
  #Extract coefficients
  coef_summary <- summary(step_model)$coefficients
  return(data.frame(Variable = rownames(coef_summary),
                    coef = coef_summary[, "coef"],
                    `exp(coef)` = coef_summary[, "exp(coef)"],
                    `se(coef)` = coef_summary[, "se(coef)"],
                    z = coef_summary[, "z"],
                    P = coef_summary[, "Pr(>|z|)"]))
}
biomarkers <- c("BCL2","COX2", "CyclinD1", "EGFRext", "HPVISHHR", "p16_binary", "PLK1", "Survivin", "TILS")
biomarkers_reduced <- c("p16_binary", "Survivin", "HPVISHHR", "TILS")
#biomarker_model <- fit_cox_model(training_data, biomarkers)
str(training_data[, c("t.os", "Outcome", biomarkers)])
sapply(training_data[, c("t.os", "Outcome", biomarkers)], function(x) sum(is.na(x)))
table(training_data$Outcome, useNA = "ifany")

colnames(training_data)[colnames(training_data) == 'Age at Diagnosis'] <- 'Age.at.Diagnosis'
colnames(training_data)[colnames(training_data) == 'Smoking status'] <- 'Smoking.status'

str(training)
n_obs <- training_data %>%
  filter(!is.na(t.os)) %>%
  nrow()

print(n_obs)

complete_training_data <- training_data[complete.cases(training_data[, c("t.os", "Outcome", all_vars1)]), ]
# n= 259, want 266
print(print_cox_model(complete_training_data, biomarkers))
#Still get EGFRext, although it shouldn't be included

print(print_cox_model(complete_training_data, biomarkers_reduced))




```
# Kaplan-Meier curves:



## Model training and classification of risk groups (+Survivin)
```{r}
# Train the biomarker-only model

fit_cox_model <- function(data, vars) {
#  data <- data[complete.cases(data[, c("t.os", "Outcome", vars)]), ]
  cat("Training n =", nrow(data), "\n")
  
  formula <- as.formula(paste("Surv(t.os, Outcome) ~", paste(vars, collapse = " + ")))
  full_model <- coxph(formula, data = data)
#  k <- qchisq(0.25, 1, lower.tail = FALSE)  # 1.323
#  step_model <- step(full_model, direction = "backward", k = k, trace = 1)
  
  return(full_model)  # Return the model object
}

# final_model (Cox) : Surv(t.os, Outcome) ~ biomarkers_reduced, with complete training data (n=259)
final_model <- fit_cox_model(complete_training_data, biomarkers_reduced)


str(val_data_dx_date[, c("t.os", "Outcome", biomarkers_reduced)])
sapply(val_data_dx_date[, c("t.os", "Outcome", biomarkers_reduced)], function(x) sum(is.na(x)))
table(val_data_dx_date$Outcome, useNA = "ifany")


colnames(val_data)[colnames(val_data) == 'Age at Diagnosis'] <- 'Age.at.Diagnosis'
colnames(val_data)[colnames(val_data) == 'Smoking status'] <- 'Smoking.status'

colnames(val_data_dx_date)[colnames(val_data_dx_date) == 'Age at Diagnosis'] <- 'Age.at.Diagnosis'
colnames(val_data_dx_date)[colnames(val_data_dx_date) == 'Smoking status'] <- 'Smoking.status'
colnames(val_data_dx_date)[colnames(val_data_dx_date) == 'Diagnostic Biopsy Date'] <- 'Diagnostic.Biopsy.Date'

complete_val_data_dx_date <- val_data_dx_date[complete.cases(val_data_dx_date[, c("t.os", "Outcome", "Diagnostic.Biopsy.Date", all_vars1)]), ]

nrow(complete_val_data_dx_date)
# n= 212

# Calculate risk scores based on training set:
complete_training_data$risk_score <- predict(final_model, complete_training_data, type = "risk")
summary(complete_training_data$risk_score)

# Median split — dichotomize into high- and low-risk groups
complete_training_data$risk_group <- ifelse(complete_training_data$risk_score > median(complete_training_data$risk_score, na.rm = TRUE), 1, 0)

sum(complete_training_data$risk_group)
risk_med_cutoff <- median(complete_training_data$risk_score)

# calculate and classify risk scores in validation cohort:
complete_val_data_dx_date$risk_score <- predict(final_model, complete_val_data_dx_date, type = "risk")
summary(complete_val_data_dx_date$risk_score)

complete_val_data_dx_date$risk_group <- ifelse(complete_val_data_dx_date$risk_score > risk_med_cutoff, 1, 0)
# Number of patients in high-risk and low-risk groups (validation set):
table(complete_val_data_dx_date$risk_group)

# No. of events in each group:
complete_val_data_dx_date %>%
  select(Outcome, risk_group) %>%
  group_by(risk_group, Outcome) %>%
  summarise(n = n())



```

## Model training and classification of risk groups (-Survivin)
```{r}
new_model <- fit_cox_model(complete_training_data, c("p16_binary", "HPVISHHR", "TILS"))

# Calculate risk scores based on median of training set:
complete_training_data$risk_score_new <- predict(new_model, complete_training_data, type = "risk")
summary(complete_training_data$risk_score_new)
summary(complete_training_data$risk_score)

ggplot(complete_training_data, aes(x = risk_score)) +
  geom_histogram(aes(y = after_stat(density), fill = "Original"), alpha = 0.5, bins = 100) +
  geom_histogram(aes(x = risk_score_new, y = after_stat(density), fill = "New"), alpha = 0.5, bins = 100) +
  labs(title = "Histogram of Risk Scores",
       x = "Risk Score",
       y = "Density",
       fill = "Score Type") +
  theme_minimal()

risk_med_cutoff_new <- median(complete_training_data$risk_score_new)

hist(complete_training_data$risk_score_new, xlab = "Risk score",
     col = "grey", border = "black", breaks = 100)
# calculate and classify risk scores in validation cohort:
complete_val_data_dx_date$risk_score_new <- predict(new_model, complete_val_data_dx_date, type = "risk")
summary(complete_val_data_dx_date$risk_score_new)
#plot
ggplot(complete_val_data_dx_date, aes(x = risk_score_new)) +
  geom_histogram(aes(y = after_stat(density), fill = "Original"), alpha = 0.5, bins = 100) +
  geom_histogram(aes(x = risk_score_new, y = after_stat(density), fill = "New"), alpha = 0.5, bins = 100) +
  labs(title = "Histogram of Risk Scores",
       x = "Risk Score",
       y = "Density",
       fill = "Score Type") +
  theme_minimal()

complete_val_data_dx_date$risk_group_new <- ifelse(complete_val_data_dx_date$risk_score_new > risk_med_cutoff_new, 1, 0)
# Number of patients in high-risk and low-risk groups (validation set):
table(complete_val_data_dx_date$risk_group_new)

# No. of events in each group:
complete_val_data_dx_date %>%
  select(Outcome, risk_group_new) %>%
  group_by(risk_group_new, Outcome) %>%
  summarise(n = n())


# 3-year and 5-year OS by risk group:

surv_obj <- Surv(complete_val_data_dx_date$t.os, complete_val_data_dx_date$Outcome)
surv_fit_new <- survfit(surv_obj ~ risk_group_new, data = complete_val_data_dx_date)
summary(surv_fit_new, times = seq(0,5)*1)


```

# classification changes in the new model
```{r}
library(dplyr)

#Validation data

complete_val_data_dx_date %>%
  group_by(risk_group_new, risk_group) %>%
  summarise(n = n())  

#Training data
# Median split: dichotomize into high- and low-risk groups
complete_training_data$risk_group_new <- ifelse(complete_training_data$risk_score_new > median(complete_training_data$risk_score_new, na.rm = TRUE), 1, 0)

complete_training_data %>%
  group_by(risk_group_new, risk_group) %>%
  summarise(n = n())


```



## Plot, 4-biomarker model
```{r}
# Survival analysis
library(survival)

# Validation data

## From documentation: Surv() creates a survival object. Takes two arguments: time to event and event status: 1 for event and 0 for censoring.
## survfit() : Fits a survival curve using either a formula, of from a previously fitted Cox model.
## We will use "surv_obj ~ risk_group" as the formula for the K-M survival curves
#3-year and 5-year OS
surv_obj <- Surv(complete_val_data_dx_date$t.os, complete_val_data_dx_date$Outcome)
surv_fit0 <- survfit(surv_obj ~ risk_group, data = complete_val_data_dx_date)
summary(surv_fit0, times = seq(0,5)*1)

# Log-rank test to conduct between-group significance -> log-rank p-value
survdiff(surv_obj ~ risk_group, data = complete_val_data_dx_date)

library(survminer)
library(ggsurvfit)

# Plot with ggsurvplot

surv_fit <- survfit2(Surv(t.os, Outcome) ~ risk_group, data = complete_val_data_dx_date, type = "kaplan-meier") %>%
  ggsurvfit() +
  add_risktable() +
  add_confidence_interval() +
  guides(fill = "none") +
  add_pvalue(caption = "Log-rank {p.value}") + #show log-rank p-value
  scale_x_continuous(
    limits = c(0,5), # Truncates at 5y
    breaks = seq(0,5, by =1 ), # 1-year ticks
    name = "Time (years)"
  ) +
  scale_y_continuous(name = "Survival Probability") +
  scale_color_manual(
    values = c("red", "blue"),
    name = "Risk Group",
    labels = c("Low Risk", "High Risk")
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_blank(),      # optional: remove grey background
    panel.grid = element_blank()             # optional: remove gridlines
  )


surv_fit


# Training data curve 

surv_fit_t <- survfit2(Surv(t.os, Outcome) ~ risk_group, data = complete_training_data, type = "kaplan-meier") %>%
  ggsurvfit() +
  add_risktable() +
  add_confidence_interval() +
  guides(fill = "none") +
  add_pvalue(caption = "Log-rank {p.value}") + #show log-rank p-value
  scale_x_continuous(
    limits = c(0,5), # Truncates at 5y
    breaks = seq(0,5, by =1 ), # 1-year ticks
    name = "Time (years)"
  ) +
  scale_y_continuous(name = "Survival Probability") +
  scale_color_manual(
    values = c("red", "blue"),
    name = "Risk Group",
    labels = c("Low Risk", "High Risk")
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_blank(),      # optional: remove grey background
    panel.grid = element_blank()             # optional: remove gridlines
  ) +
  annotate(
    "text", 
    x = 0,  # adjust X location as needed
    y = 0.05,  # adjust Y location as needed
    label = "HR = 8.20 (4.19-16.05)", 
    hjust = 0,
    size = 4
  )

surv_fit_t

```

#Plots, 3-biomarker model
```{r}

# Validation data curve 

surv_fit_new <- survfit2(Surv(t.os, Outcome) ~ risk_group_new, data = complete_val_data_dx_date, type = "kaplan-meier") %>%
  ggsurvfit() +
  add_risktable() +
  add_confidence_interval() +
  guides(fill = "none") +
  add_pvalue(caption = "Log-rank {p.value}") + #show log-rank p-value
  scale_x_continuous(
    limits = c(0,5), # Truncates at 5y
    breaks = seq(0,5, by =1 ), # 1-year ticks
    name = "Time (years)"
  ) +
  scale_y_continuous(name = "Survival Probability") +
  scale_color_manual(
    values = c("red", "blue"),
    name = "Risk Group",
    labels = c("Low Risk", "High Risk")
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_blank(),      # optional: remove grey background
    panel.grid = element_blank()             # optional: remove gridlines
  )

surv_fit_new


# Training data curve 

surv_fit_new_t <- survfit2(Surv(t.os, Outcome) ~ risk_group_new, data = complete_training_data, type = "kaplan-meier") %>%
  ggsurvfit() +
  add_risktable() +
  add_confidence_interval() +
  guides(fill = "none") +
  add_pvalue(caption = "Log-rank {p.value}") + #show log-rank p-value
  scale_x_continuous(
    limits = c(0,5), # Truncates at 5y
    breaks = seq(0,5, by =1 ), # 1-year ticks
    name = "Time (years)"
  ) +
  scale_y_continuous(name = "Survival Probability") +
  scale_color_manual(
    values = c("red", "blue"),
    name = "Risk Group",
    labels = c("Low Risk", "High Risk")
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_blank(),      # optional: remove grey background
    panel.grid = element_blank()             # optional: remove gridlines
  ) +
  annotate(
    "text", 
    x = 0,  # adjust X location as needed
    y = 0.05,  # adjust Y location as needed
    label = "HR = 5.69 (3.38–9.58)", 
    hjust = 0,
    size = 4
  )

surv_fit_new_t
```





## Table S11 (includes C-index, sensitivity, PPV, NPV)
```{r}
# Train the biomarker-only model

fit_cox_model <- function(data, vars) {
#  data <- data[complete.cases(data[, c("t.os", "Outcome", vars)]), ]
  cat("Training n =", nrow(data), "\n")
  
  formula <- as.formula(paste("Surv(t.os, Outcome) ~", paste(vars, collapse = " + ")))
  full_model <- coxph(formula, data = data)
#  k <- qchisq(0.25, 1, lower.tail = FALSE)  # 1.323
#  step_model <- step(full_model, direction = "backward", k = k, trace = 1)
  
  return(full_model)  # Return the model object
}

final_model <- fit_cox_model(complete_training_data, biomarkers_reduced)


str(val_data_dx_date[, c("t.os", "Outcome", biomarkers_reduced)])
sapply(val_data_dx_date[, c("t.os", "Outcome", biomarkers_reduced)], function(x) sum(is.na(x)))
table(val_data_dx_date$Outcome, useNA = "ifany")


colnames(val_data)[colnames(val_data) == 'Age at Diagnosis'] <- 'Age.at.Diagnosis'
colnames(val_data)[colnames(val_data) == 'Smoking status'] <- 'Smoking.status'

colnames(val_data_dx_date)[colnames(val_data_dx_date) == 'Age at Diagnosis'] <- 'Age.at.Diagnosis'
colnames(val_data_dx_date)[colnames(val_data_dx_date) == 'Smoking status'] <- 'Smoking.status'
colnames(val_data_dx_date)[colnames(val_data_dx_date) == 'Diagnostic Biopsy Date'] <- 'Diagnostic.Biopsy.Date'

complete_val_data_dx_date <- val_data_dx_date[complete.cases(val_data_dx_date[, c("t.os", "Outcome", "Diagnostic.Biopsy.Date", all_vars1)]), ]

nrow(complete_val_data_dx_date)
# n= 212

complete_val_data_dx_date$risk_score <- predict(final_model, newdata = complete_val_data_dx_date, type = "lp")
summary(complete_val_data_dx_date$risk_score)

# Median split: dichotomize into high- and low-risk groups
complete_val_data_dx_date$risk_group <- ifelse(complete_val_data_dx_date$risk_score > median(complete_val_data_dx_date$risk_score, na.rm = TRUE), 1, 0)

# Survival analysis
library(survival)
surv_obj <- Surv(complete_val_data_dx_date$t.os, complete_val_data_dx_date$Outcome)
surv_fit <- survfit(surv_obj ~ risk_group, data = complete_val_data_dx_date)
summary(surv_fit)
survdiff(surv_obj ~ risk_group, data = complete_val_data_dx_date)  # Log-rank test

plot(surv_fit, col = c("blue", "red"), main = "Survival by Risk Group", xlab = "Time (t.os)", ylab = "Survival Probability", xlim = c(0,5))
legend("topright", c("Low Risk", "High Risk"), col = c("blue", "red"), lty = 1)

library(survminer)

# Fit survival model
surv_fit <- survfit(Surv(t.os, Outcome) ~ risk_group, data = complete_val_data_dx_date, type = "kaplan-meier")

# Plot with ggsurvplot
ggsurvplot(
  surv_fit,
  data = complete_val_data_dx_date,
  xlim = c(0, 5),                   # restrict x-axis to 5 years
  break.time.by = 1,
  xlab = "Time (years)",
  ylab = "Survival Probability",
 #surv.median.line = "hv",         # optional: add horizontal/vertical median lines
  pval = TRUE,                     # show log-rank p-value
  conf.int = TRUE,                 # add confidence interval
  risk.table = TRUE,              # add risk table below
  tables.y.text = FALSE,
  censor = TRUE,                   # mark censoring (default: TRUE)
  mark.time = TRUE,                # show tick marks at event times
  palette = c("blue", "red"),      # customise colours
  legend.title = "Risk Group",
  legend.labs = c("Low Risk", "High Risk"),
  fontsize = 3
)

surv_fit
```



# Cox model, 4-biomarker model
```{r}
# Cox model for HR, CI, p-value for validation data — 4-biomarker model

cox_model <- coxph(Surv(t.os, Outcome) ~ risk_group, data = complete_val_data_dx_date)
cox_summary <- summary(cox_model)
print(cox_summary$coefficients)  # See row/col names
coef <- cox_summary$coefficients["risk_group", "coef"]
hr <- cox_summary$coefficients["risk_group", "exp(coef)"]
se <- cox_summary$coefficients["risk_group", "se(coef)"]
ci_lower <- exp(coef - 1.96 * se)
ci_upper <- exp(coef + 1.96 * se)
p_value <- cox_summary$coefficients["risk_group", "Pr(>|z|)"]

# Output
cat("HR (1 vs. 0):", hr, "\n") #HR (1 vs. 0): 4.24
cat("95% CI:", ci_lower, "-", ci_upper, "\n") #95% CI: 2.49 - 7.23 
cat("P-value:", p_value, "\n") #P-value: p<0.001 (1.1e-7) 

#More succinct way to find HR using val data

library(gtsummary)

coxph(Surv(t.os, Outcome) ~ risk_group, data = complete_val_data_dx_date) %>%
  tbl_regression(exp=TRUE)

# HR using val data, 4-biomarker

hr1 <- hazard.ratio(x=complete_val_data_dx_date$risk_group, surv.time = complete_val_data_dx_date$t.os, surv.event = complete_val_data_dx_date$Outcome)
hr1[c(1,4,5,6)]

#HR using training data:

complete_training_data$risk_score <- predict(final_model, newdata = complete_training_data, type = "risk")

# Median split: dichotomize into high- and low-risk groups
complete_training_data$risk_group <- ifelse(complete_training_data$risk_score > median(complete_training_data$risk_score, na.rm = TRUE), 1, 0)

hr_train <- hazard.ratio(x=complete_training_data$risk_group, surv.time = complete_training_data$t.os, surv.event = complete_training_data$Outcome)

# HR (95%CI), p-val
hr_train[c(1,4,5,6)]
# Get 8.21 (4.19-16.05)

cox_risk_analysis <- function(data, time_col = "t.os", event_col = "Outcome", group_col = "risk_group", factor = FALSE) {
  # Handle risk_group
  if (factor) {
    data[[group_col]] <- factor(data[[group_col]], levels = c("Low", "High"))
    coef_name <- paste0(group_col, "High")
  } else {
    data[[group_col]] <- as.numeric(data[[group_col]])
    coef_name <- group_col
  }
  
  # Cox model
  formula <- as.formula(paste("Surv(", time_col, ",", event_col, ") ~", group_col))
  cox_model <- coxph(formula, data = data)
  cox_summary <- summary(cox_model)  
  
  # Extract results
  coef <- cox_summary$coefficients[coef_name, "coef"]
  hr <- cox_summary$coefficients[coef_name, "exp(coef)"]
  se <- cox_summary$coefficients[coef_name, "se(coef)"]
  ci_lower <- exp(coef - 1.96 * se)
  ci_upper <- exp(coef + 1.96 * se)
  p_value <- cox_summary$coefficients[coef_name, "Pr(>|z|)"]
  # Output
  label <- if (factor) "High vs. Low" else "1 vs. 0"
  cat("HR (", label, "): ", hr, "\n", sep = "")
  cat("95% CI: ", ci_lower, " - ", ci_upper, "\n", sep = "")
  cat("P-value: ", p_value, "\n", sep = "")
  
  # Return
  return(list(hr = hr, ci_lower = ci_lower, ci_upper = ci_upper, p_value = p_value, 
              coefficients = cox_summary$coefficients))
}

results_val <- cox_risk_analysis(complete_val_data_dx_date, time_col = "t.os", 
                                   event_col = "Outcome", group_col = "risk_group", 
                                   factor = FALSE)


```


# Cox model, 3-biomarker model
```{r}
# Cox model for HR, CI, p-value for validation data — 3-biomarker model

#Risk scores
complete_val_data_dx_date$risk_score_new <- predict(new_model, newdata = complete_val_data_dx_date, type = "risk")

# Median split: dichotomize into high- and low-risk groups
complete_val_data_dx_date$risk_group_new <- ifelse(complete_val_data_dx_date$risk_score_new > median(complete_training_data$risk_score_new, na.rm = TRUE), 1, 0)


cox_model_new <- coxph(Surv(t.os, Outcome) ~ risk_group_new, data = complete_val_data_dx_date)

cox_summary_new <- summary(cox_model_new)
print(cox_summary_new$coefficients)  # See row/col names
coef_new <- cox_summary_new$coefficients["risk_group_new", "coef"]
hr_new <- cox_summary_new$coefficients["risk_group_new", "exp(coef)"]
se_new <- cox_summary_new$coefficients["risk_group_new", "se(coef)"]
ci_lower_new <- exp(coef_new - 1.96 * se_new)
ci_upper_new <- exp(coef_new + 1.96 * se_new)
p_value_new <- cox_summary_new$coefficients["risk_group_new", "Pr(>|z|)"]

# Output
cat("HR (1 vs. 0):", hr_new, "\n") #HR (1 vs. 0): 3.82
cat("95% CI:", ci_lower_new, "-", ci_upper_new, "\n") #95% CI: 2.41 - 6.07
cat("P-value:", p_value_new, "\n") #P-value: p<0.001 (1.3e-8) 

# For a table of HR, 95% CI using val data 

library(gtsummary)

coxph(Surv(t.os, Outcome) ~ risk_group_new, data = complete_val_data_dx_date) %>%
  tbl_regression(exp=TRUE)


hr1_new <- hazard.ratio(x=complete_val_data_dx_date$risk_group_new, surv.time = complete_val_data_dx_date$t.os, surv.event = complete_val_data_dx_date$Outcome)
hr1_new[c(1,4,5,6)]


hr_train_new <- hazard.ratio(x=complete_training_data$risk_group_new, surv.time = complete_training_data$t.os, surv.event = complete_training_data$Outcome)

# HR (95%CI), p-val
hr_train_new[c(1,4,5,6)]
```

# HR comparisons
```{r}
# NOTE: COMPARING HR IS NOT THE BEST WAY TO COMPARE MODELS. HRs tell you how quickly disease progresses on average but not discrimination.
# Comparing HRs between high- and low-risk groups in validation data, 4-biomarker model vs. 3-biomarker model
hr.comp(hr1, hr1_new)
# And for training data, 4-biomarker model vs. 3-biomarker model:
hr.comp(hr_train, hr_train_new)

```


# Discrimination Measures: C-index
```{r}
library(pacman)
pacman::p_load(survival,
               Hmisc,
               riskRegression,
               timeROC)


summary(complete_val_data_dx_date$t.os)
# Distribution of time to OS 
tOS_hist <- hist(complete_val_data_dx_date$t.os)

#Add linear predictor for 4-biomarker model (lp4) and 3-biomarker model (lp3) to the validation set
complete_val_data_dx_date$lp4 <- predict(final_model, complete_val_data_dx_date)
complete_val_data_dx_date$lp3 <- predict(new_model, complete_val_data_dx_date)


# FOUR-BIOMARKER (ORIGINAL) CONCORDANCE, VALIDATION DATA
# 3-year C-index:
## Harrell's C
cindex_3yr <- complete_val_data_dx_date %>%
  filter(t.os <= 3) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp4, data = ., reverse = TRUE) }

## Uno's C
uno_cindex_3yr <- complete_val_data_dx_date %>%
  filter(t.os <= 3) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp4, data = ., reverse = TRUE,
                           timewt = "n/G2") }

# 5-year C-index:
## Harrell's C
cindex_5yr <- complete_val_data_dx_date %>%
  filter(t.os <= 5) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp4, data = ., reverse = TRUE) }

## Uno's C
uno_cindex_5yr <- complete_val_data_dx_date %>%
  filter(t.os <= 5) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp4, data = ., reverse = TRUE,
                           timewt = "n/G2") }

# 10-year C-index:
## Harrell's C
cindex_10yr <- complete_val_data_dx_date %>%
  filter(t.os <= 10) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp4, data = ., reverse = TRUE) }

## Uno's C
uno_cindex_10yr <- complete_val_data_dx_date %>%
  filter(t.os <= 10) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp4, data = ., reverse = TRUE,
                           timewt = "n/G2") }


## Validation data: calculation of C-index for 3-biomarker model

# THREE-BIOMARKER (NEW) CONCORDANCE, VALIDATION DATA
# 3-year C-index:
## Harrell's C
cindex_3yr_new <- complete_val_data_dx_date %>%
  filter(t.os <= 3) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp3, data = ., reverse = TRUE) }

## Uno's C
uno_cindex_3yr_new <- complete_val_data_dx_date %>%
  filter(t.os <= 3) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp3, data = ., reverse = TRUE,
                           timewt = "n/G2") }

# 5-year C-index:
## Harrell's C
cindex_5yr_new <- complete_val_data_dx_date %>%
  filter(t.os <= 5) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp3, data = ., reverse = TRUE) }

## Uno's C
uno_cindex_5yr_new <- complete_val_data_dx_date %>%
  filter(t.os <= 5) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp3, data = ., reverse = TRUE,
                           timewt = "n/G2") }

# 10-year C-index:
## Harrell's C
cindex_10yr_new <- complete_val_data_dx_date %>%
  filter(t.os <= 10) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp3, data = ., reverse = TRUE) }

## Uno's C
uno_cindex_10yr_new <- complete_val_data_dx_date %>%
  filter(t.os <= 10) %>%  # keep events or censored before 3 years
  { concordance(Surv(t.os, Outcome) ~ lp3, data = ., reverse = TRUE,
                           timewt = "n/G2") }
```


# Concordance indices
```{r, concordance_table,warning=FALSE,echo=FALSE}

alpha <- 0.05

temp <- c(
  # Original model
  ## 3-year Harrell's C
  cindex_3yr$concordance,
  cindex_3yr$concordance - 
    qnorm(1 - alpha/2) * sqrt(cindex_3yr$var),
  cindex_3yr$concordance + 
    qnorm(1 - alpha/2) * sqrt(cindex_3yr$var),
  ## 5-year Harrell's C
  cindex_5yr$concordance,
  cindex_5yr$concordance - 
    qnorm(1 - alpha/2) * sqrt(cindex_5yr$var),
  cindex_5yr$concordance + 
    qnorm(1 - alpha/2) * sqrt(cindex_5yr$var),
  ## 10-year Harrell's C
  cindex_10yr$concordance,
  cindex_10yr$concordance - 
    qnorm(1 - alpha/2) * sqrt(cindex_10yr$var),
  cindex_10yr$concordance + 
    qnorm(1 - alpha/2) * sqrt(cindex_10yr$var),
  
  ## 3-year Uno's C
  uno_cindex_3yr$concordance,
  uno_cindex_3yr$concordance - 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_3yr$var),
  uno_cindex_3yr$concordance + 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_3yr$var),
  ## 5-year Uno's C
  uno_cindex_5yr$concordance,
  uno_cindex_5yr$concordance - 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_5yr$var),
  uno_cindex_5yr$concordance + 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_5yr$var),
  ## 10-year Uno's C
  uno_cindex_10yr$concordance,
  uno_cindex_10yr$concordance - 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_10yr$var),
  uno_cindex_10yr$concordance + 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_10yr$var),
  
  # New model
  ## 3-year Harrell's C
  cindex_3yr_new$concordance,
  cindex_3yr_new$concordance - 
    qnorm(1 - alpha/2) * sqrt(cindex_3yr_new$var),
  cindex_3yr_new$concordance + 
    qnorm(1 - alpha/2) * sqrt(cindex_3yr_new$var),
  ## 5-year Harrell's C
  cindex_5yr_new$concordance,
  cindex_5yr_new$concordance - 
    qnorm(1 - alpha/2) * sqrt(cindex_5yr_new$var),
  cindex_5yr_new$concordance + 
    qnorm(1 - alpha/2) * sqrt(cindex_5yr_new$var),
  ## 10-year Harrell's C
  cindex_10yr_new$concordance,
  cindex_10yr_new$concordance - 
    qnorm(1 - alpha/2) * sqrt(cindex_10yr_new$var),
  cindex_10yr_new$concordance + 
    qnorm(1 - alpha/2) * sqrt(cindex_10yr_new$var),
  
  ## 3-year Uno's C
  uno_cindex_3yr_new$concordance,
  uno_cindex_3yr_new$concordance - 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_3yr_new$var),
  uno_cindex_3yr_new$concordance + 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_3yr_new$var),
  ## 5-year Uno's C
  uno_cindex_5yr_new$concordance,
  uno_cindex_5yr_new$concordance - 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_5yr_new$var),
  uno_cindex_5yr_new$concordance + 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_5yr_new$var),
  ## 10-year Uno's C
  uno_cindex_10yr_new$concordance,
  uno_cindex_10yr_new$concordance - 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_10yr_new$var),
  uno_cindex_10yr_new$concordance + 
    qnorm(1 - alpha/2) * sqrt(uno_cindex_10yr_new$var)


)


res_C  <- matrix(temp, 
                nrow = 2, 
                ncol = 18, 
                byrow = TRUE,
                dimnames = list(
  c("Four-biomarker model", 
    "Three-biomarker model"),
  
  c("Harrell - 3y", "Lower .95", "Upper .95",
    "Harrell - 5y", "Lower .95", "Upper .95",
    "Harrell - 10y", "Lower .95", "Upper .95",
    "Uno - 3y", "Lower .95", "Upper .95",
    "Uno - 5y", "Lower .95", "Upper .95",
    "Uno - 10y", "Lower .95", "Upper .95"
    ))
)

res_C <- round(res_C, 3) # Digit
res_C
```


# PPV, NPV, sensitivity

```{r}

# Four-biomarker model
# contingency table for PPV, NPV, sensitivity calculations
table_pred <- table(Predicted = complete_val_data_dx_date$risk_group, Actual = complete_val_data_dx_date$Outcome)
TP <- table_pred[2,2]; FP <- table_pred[2,1]; TN <- table_pred[1,1]; FN <- table_pred[1,2]
print(sens <- TP/(TP+FN))
print(PPV <- TP/(TP+FP))
print(NPV <- TN/(TN+FN))

# Three-biomarker model
# contingency table for PPV, NPV, sensitivity calculations
table_pred <- table(Predicted = complete_val_data_dx_date$risk_group_new, Actual = complete_val_data_dx_date$Outcome)
TP <- table_pred[2,2]; FP <- table_pred[2,1]; TN <- table_pred[1,1]; FN <- table_pred[1,2]
print(sens <- TP/(TP+FN))
print(PPV <- TP/(TP+FP))
print(NPV <- TN/(TN+FN))

table_pred

```
## Surgery vs no surgery in high-risk and low-risk patient groups, validation cohort
```{r}
# summary(complete_val_data$Surgery) #total: 212. Paper (Fig 3C): 221
# table_risk_surgery <- table(Risk_group = complete_val_data$risk_group, Surgery = complete_val_data$Surgery); print(table_risk_surgery)
# ls0 <- table_risk_surgery[1,1]; ls1 <- table_risk_surgery[1,2]; hs0 <- table_risk_surgery[2,1]; hs1 <- table_risk_surgery[2,2]
# 
# #Split into the risk+surgery groups:
# val_data_ls0 <- complete_val_data[(complete_val_data$Surgery == 0 & complete_val_data$risk_group == 0),]
# val_data_ls1 <- complete_val_data[(complete_val_data$Surgery == 1 & complete_val_data$risk_group == 0),]
# val_data_hs0 <- complete_val_data[(complete_val_data$Surgery == 0 & complete_val_data$risk_group == 1),]
# val_data_hs1 <- complete_val_data[(complete_val_data$Surgery == 1 & complete_val_data$risk_group == 1),]
# 
# table(complete_val_data$risk_group)
# #n = 212 (vs. 221)
# #Low: 95 (vs. 101)
# #High: 117 (vs. 120)
# 
# nrow(val_data_ls0) #39, not 38
# nrow(val_data_ls1) #67, not 63
# 
# nrow(val_data_hs0) # 48, not 55
# nrow(val_data_hs1) # 58, not 65
# 
# table(val_data_ls0$Smoking.status)
# table(val_data_ls1$Smoking.status)
# table(val_data_ls0$HPVISHHR)
# table(val_data_ls1$HPVISHHR)
# 
# table(val_data_hs0$Smoking.status)
# table(val_data_hs1$Smoking.status)
# table(val_data_hs0$HPVISHHR)
# table(val_data_hs1$HPVISHHR)
# 
# 
# library(survminer)
# 
# #Create combined group factor
# complete_val_data$risk_surgery <- factor(
#   paste0(
#     
#     ifelse(complete_val_data$risk_group==0, "l", "h"),
#     ifelse(complete_val_data$Surgery == 0, "s0", "s1")
#     
#   ),
#   levels = c("ls0", "ls1", "hs0", "hs1")
# )
# 
# #Kaplan-Meier
# surv_obj <- Surv(complete_val_data$t.os, complete_val_data$Outcome)
# 
# surv_fit <- survfit(surv_obj ~ risk_surgery, data = complete_val_data)
# 
# str(complete_val_data$risk_surgery)
# print(summary(surv_fit))
# 
# ggsurv <- ggsurvplot(
#   surv_fit,
#   data = complete_val_data,
#   palette = c("blue", "green", "red", "purple"),
#   risk.table = TRUE,
#   risk.table.height = 0.25,
#   title = "Survival by Risk Group and Surgery Status",
#   xlab = "Time (months)",
#   ylab = "Survival Probability",
#   legend.title = "Group",
#   legend.labs = c("Low, Surgery-", "Low, Surgery+", "High, Surgery-", "High, Surgery+"),
#   conf.int = FALSE,
#   pval = FALSE,
#   xlim = c(0, 60)  # Adjust based on t.os range
# )
# print(ggsurv)
# 
# 
# low_data <- complete_val_data[complete_val_data$risk_group == 0, ]
# low_cox <- coxph(Surv(t.os, Outcome) ~ Surgery, data = low_data)
# low_summary <- summary(low_cox)
# low_summary
# low_hr <- low_summary$coefficients["Surgery1", "exp(coef)"]
# low_ci <- exp(low_summary$coefficients["Surgery1", "coef"] + c(-1.96, 1.96) * low_summary$coefficients["Surgery1", "se(coef)"])
# low_p <- low_summary$coefficients["Surgery1", "Pr(>|z|)"]
# 
# low_hr
# low_ci
# low_p
# 
# 
# high_data <- complete_val_data[complete_val_data$risk_group == 1, ]
# high_cox <- coxph(Surv(t.os, Outcome) ~ Surgery, data = high_data)
# high_summary <- summary(high_cox)
# high_summary
# high_hr <- high_summary$coefficients["Surgery1", "exp(coef)"]
# high_ci <- exp(high_summary$coefficients["Surgery1", "coef"] + c(-1.96, 1.96) * high_summary$coefficients["Surgery1", "se(coef)"])
# high_p <- high_summary$coefficients["Surgery1", "Pr(>|z|)"]
# 
# #High-risk group: surgery vs non surgery
# high_hr
# high_ci
# high_p
# 
# 
# ```
# ## Surgery vs no surgery in low-risk and high-risk pts, training cohort
# ```{r}
# summary(complete_training_data$Surgery) #total: 259 Paper (Fig 3A): 266
# table_risk_surgery0 <- table(Risk_group = complete_training_data$risk_group, Surgery = complete_training_data$Surgery); print(table_risk_surgery0)
# ls0 <- table_risk_surgery0[1,1]; ls1 <- table_risk_surgery0[1,2]; hs0 <- table_risk_surgery0[2,1]; hs1 <- table_risk_surgery0[2,2]
# 
# #Split into the risk+surgery groups:
# train_data_ls0 <- complete_training_data[(complete_training_data$Surgery == 0 & complete_training_data$risk_group == 0),]
# train_data_ls1 <- complete_training_data[(complete_training_data$Surgery == 1 & complete_training_data$risk_group == 0),]
# train_data_hs0 <- complete_training_data[(complete_training_data$Surgery == 0 & complete_training_data$risk_group == 1),]
# train_data_hs1 <- complete_training_data[(complete_training_data$Surgery == 1 & complete_training_data$risk_group == 1),]
# 
# 
# library(survminer)
# 
# #Create combined group factor
# complete_training_data$risk_surgery <- factor(
#   paste0(
#     
#     ifelse(complete_training_data$risk_group==0, "l", "h"),
#     ifelse(complete_training_data$Surgery == 0, "s0", "s1")
#     
#   ),
#   levels = c("ls0", "ls1", "hs0", "hs1")
# )
# 
# #Kaplan-Meier
# surv_obj0 <- Surv(complete_training_data$t.os, complete_training_data$Outcome)
# 
# surv_fit0 <- survfit(surv_obj0 ~ risk_surgery, data = complete_training_data)
# 
# str(complete_training_data$risk_surgery)
# print(summary(surv_fit0))
# 
# ggsurv <- ggsurvplot(
#   surv_fit0,
#   data = complete_training_data,
#   palette = c("blue", "green", "red", "purple"),
#   risk.table = TRUE,
#   risk.table.height = 0.25,
#   title = "Survival by Risk Group and Surgery Status",
#   xlab = "Time (months)",
#   ylab = "Survival Probability",
#   legend.title = "Group",
#   legend.labs = c("Low, Surgery-", "Low, Surgery+", "High, Surgery-", "High, Surgery+"),
#   conf.int = FALSE,
#   pval = FALSE,
#   xlim = c(0, 60)  # Adjust based on t.os range
# )
# print(ggsurv)
# 
# 
# low_data <- complete_training_data[complete_training_data$risk_group == 0, ]
# low_cox <- coxph(Surv(t.os, Outcome) ~ Surgery, data = low_data)
# low_summary <- summary(low_cox)
# low_summary
# low_hr <- low_summary$coefficients["Surgery1", "exp(coef)"]
# low_ci <- exp(low_summary$coefficients["Surgery1", "coef"] + c(-1.96, 1.96) * low_summary$coefficients["Surgery1", "se(coef)"])
# low_p <- low_summary$coefficients["Surgery1", "Pr(>|z|)"]
# 
# low_hr
# low_ci
# low_p
# 
# 
# high_data <- complete_training_data[complete_training_data$risk_group == 1, ]
# high_cox <- coxph(Surv(t.os, Outcome) ~ Surgery, data = high_data)
# high_summary <- summary(high_cox)
# high_summary
# high_hr <- high_summary$coefficients["Surgery1", "exp(coef)"]
# high_ci <- exp(high_summary$coefficients["Surgery1", "coef"] + c(-1.96, 1.96) * high_summary$coefficients["Surgery1", "se(coef)"])
# high_p <- high_summary$coefficients["Surgery1", "Pr(>|z|)"]
# 
# #High-risk group: surgery vs non surgery
# high_hr
# high_ci
# high_p

```


